Viscous Transport in Eroding Porous Media

### 2019-10-01 docs/transport/erosion_transport.tex
NM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations}
\label{sec:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by defining the main variables used to model erosion.  We only
briefly summarize the model, and a more detailed description can be
found in previous work~\cite{qua-moo2018}.  We consider flows inside a
confined geometry $\Omega$ that contains $M$ eroding bodies with
boundaries $\gamma_\ell$, $\ell = 1,\ldots,M$.  The boundary of the
fluid domain is $\bd \Omega = \Gamma \cup \gamma_1 \cup \cdots \cup
\gamma_M$, where $\Gamma$ is the outer boundary, taken to be a slightly
smoothed version of $[-3,3] \times [-1,1]$.  We only place eroding
bodies in $[-1,1] \times [-1,1]$ so as to create a buffer region for the
flow profile imposed at the inlet to transition to the more complex flow
intervening between the bodies. On each eroding body, $\nn$ is the
normal vector pointing into the body, and $\ss$ is the unit tangent
vector pointing in the counterclockwise direction.  The normal vector on
$\Gamma$ is the outward unit normal.  Neglecting inertial forces, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}, \\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass}, \\
    \uu = \mathbf{0}, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the eroding bodies}, \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity}, \\
    \Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
      &&\mbox{\em erosion model}.
  \end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, $\Vn$ is the normal velocity
of $\gamma$, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on $\gamma$. We simulate erosion by alternating
between solving the fluid equations \nick{and} advancing the grains according to
the erosion rate for a small time step. The strength of $\UU$ is
adjusted at each time step to achieve a constant pressure drop across
the channel.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Double-Layer Potential Formulation in $\RR^2$}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Applying the same approach as our previous work~\cite{qua-moo2018}, we
write the velocity $\uu$ as a double-layer potential 
\begin{align}
  \DDD[\eeta](\xx) = \int_{\bd\Omega} D(\xx,\yy) \eeta(\yy)\, ds_\yy = 
  \frac{1}{\pi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:velocityDLP}
\end{align}
where $D$ is the kernel of the integral operator, $\rr = \xx - \yy$,
$\rho = \|\rr\|$, $\nn$ is the unit outward normal at $\yy$, and $\eeta$
is an unknown density function.  The double-layer potential can not
represent all solutions of the incompressible Stokes equations, so we
complete the integral equation formulation by adding the $M$ Stokeslets
and rotlets~\cite{pow-mir1987}
\begin{align}
  S[\llambda_\ell,\cc_\ell](\xx) = \frac{1}{4\pi} \left( 
    -\log \rho \II + \frac{\rr \otimes \rr}{\rho^2}
    \right)\llambda_\ell, \quad
  R[\xi_\ell,\cc_\ell](\xx) = \frac{\rr^\perp}{\rho^2} \xi_\ell,
  \quad \ell=1,\ldots,M,
\end{align}
where $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ eroding body, $\rr^\perp = (r_2,-r_1)$, and $\rho =
\|\rr\|$.  Then, for any sufficiently smooth geometry $\Omega$, the
solution of the incompressible Stokes equation with a Dirichlet boundary
condition $\ff$ is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:BIE}
\begin{alignat}{3}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx) +
    \NN_0[\eeta](\xx), \quad &&\xx \in \bd\Omega, \\
  \llambda_\ell &= \frac{1}{2\pi} \int_{\gamma_\ell} 
    \eeta(\yy)\, ds_\yy &&\ell = 1,\ldots,M, \\
  \xi_\ell &= \frac{1}{2\pi} \int_{\gamma_\ell}
    (\yy - \cc_\ell)^\perp \cdot \eeta(\yy)\, ds_\yy 
    &&\ell = 1,\ldots,M.
\end{alignat}
\end{subequations}
Here, the null space associated with the flux-free condition of $\ff$ is
addressed with the integral operator $\NN_0$ which is the integral
operator with kernel $N_0(\xx,\yy) = \nn(\xx) \otimes \nn(\yy)$,
$\xx,\yy \in \Gamma$.  In this work, $\ff$ is the prescribed velocity,
$\UU$, on the outer wall, $\Gamma$, and is zero on the eroding bodies,
$\gamma_\ell$, $\ell=1,\ldots,M$.

Once~\eqref{eqn:BIE} is solved for the density function $\eeta$, the
Stokeslets $\llambda_\ell$, and the rotlets $\xi_\ell$, we are able to
compute the deformation tensor, pressure, and vorticity, defined
respectively as
\begin{align}
  \label{eqn:deformationDLP} 
  \ssigma(\xx) &= \int_{\bd\Omega} \mathbf{D}_{\ssigma}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = \frac{1}{2\pi}\int_{\bd\Omega} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \nonumber \\
    & \qquad\qquad\qquad\qquad\qquad\qquad \left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
    \right) \, ds_\yy, \\
%
  p(\xx) &= \int_{\bd\Omega} D_{p}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\pi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) \nn \cdot
      \eeta(\yy)\,ds_\yy, \\
%
  \label{eqn:vorticityDLP} 
  \omega(\xx) &= \int_{\bd\Omega} D_{\omega}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\pi} \int_{\bd\Omega}
    \frac{(\rr \cdot \nn^\perp) + (\rr \cdot \nn)}{\rho^4} 
      (\rr \cdot \eeta) \, ds_\yy.
\end{align}
Since the deformation tensor of the double-layer potential is
discontinuous across $\bd\Omega$, the jump term 
\begin{align}
  \frac{1}{2} \left(\pderiv{\eeta}{\ss} \cdot \ss \right) \left[
    \begin{array}{cc}
      s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
    \end{array}
  \right],
  \label{eqn:deformationJump}
\end{align}
must be added to $\ssigma(\xx)$ for $\xx \in \bd\Omega$.  Finally, the
Stokeslets and rotlets do not exhibit any jumps across $\gamma$, so the
associated components of the deformation tensor, pressure, and vorticity
due to the Stokeslets and rotlets are straightforward to
compute~\cite{poz1992}.  Having computed the deformation tensor on
$\gamma$, the shear stress is computed as defined in
equation~\eqref{eqn:shearStress}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In equations~\eqref{eqn:velocityDLP} and~\eqref{eqn:deformationDLP},
$\uu$ and its deformation tensor are written as layer potentials in
$\RR^2$, but the quadrature method we introduce in
Section~\ref{sec:method} requires a complex-valued representation of the
layer potentials. The first step to form a complex representation is to
write the Laplace double-layer potential as the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Re (v(x)) \quad 
    \text{where} \quad v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy, \quad x \in \Omega,
  \label{eqn:laplaceComplex}
\end{align}
where $x,y \in \CC$ are the complex counterparts of $\xx,\yy \in \RR^2$,
and $\eta$ is the complex counterpart of $\eeta$.
Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy integral by
first finding the boundary data of $v$.  Assuming that $\Omega$ is a
simply-connected interior domain, the boundary data of $v$ satisfies the
Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \bd\Omega.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
decomposed into its different connected components and the appropriate
jump relation is applied.  Since $v$ is a holomorphic function and we
have its boundary data, by the Cauchy integral theorem we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
\end{align}
for $x \in \Omega$.  Since $v(x)$ depends on the complex-valued density
function $\eta$, we use the notation $v[\eta](x)$ for the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation for the Gradient of the
Double-Layer Potential}
\label{sec:gradDLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Computing the shear stress and vorticity requires a complex-valued layer
potential representation of the gradient of the velocity.  The
deformation tensor at $x \in \Omega$ is found by computing the
derivatives of the expressions for $u_1$ and $u_2$ in
equation~\eqref{eqn:cauchyVelocity}  
\begin{equation}
\label{eqn:cauchyGradient}
  \begin{aligned}
    \pderiv{u_1}{x_1} &= +\Re (v'[\psi_1](x)) + 
      \Re (v''[y\cdot\eta](x)) - \Re (v'[\eta_1](x)) - 
      x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x)), \\
    \pderiv{u_1}{x_2} &= - \Im (v'[\psi_1](x)) - 
      \Im (v''[y\cdot\eta](x)) + x_1\Im (v''[\eta_1](x)) - 
      \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_1} &= +\Re (v'[\psi_2](x)) - 
      \Im (v''[y\cdot\eta](x)) + \Im (v'[\eta_1](x)) +
      x_1\Im (v''[\eta_1](x)) + x_2\Im (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_2} &= -\Im (v'[\psi_2](x)) - 
      \Re (v''[y\cdot\eta](x)) + x_1\Re (v''[\eta_1](x)) +
      \Im (v'[\eta_2](x)) + x_2\Re (v''[\eta_2](x)).
  \end{aligned}
\end{equation}
The same expressions are used to compute the deformation tensor for $x
\in \bd\Omega$, except that the jump
condition~\eqref{eqn:deformationJump} is included.  Finally, to compute
the shear stress, the deformation tensor on $\bd\Omega$ is applied to
the normal and tangent vectors as defined in
equation~\eqref{eqn:shearStress}.  The gradient of the velocity field is
also used to compute the vorticity in the fluid bulk.  For $x \in
\Omega$, the Cauchy integral representation of the vorticity is
\begin{align}
  \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} = 
\Re (v'[\psi_2](x)) + \Im (v'[\psi_1](x)) 
 + \Re (v'[\eta_2](x))+ \Im (v'[\eta_1](x)), \quad x \in \Omega.
\end{align}

@@

To establish an anomalous dispersion rate, the trajectories must pass
several grains so that an asymptotic regime is reached. The geometries
that we consider are too short to observe the asymptotic dispersion, so
we use a reinsertion method to form longer trajectories. Similar to the
work of others~\cite{dea-qua-bir-jua2018, puy-gou-den2019}, once a
particle reaches the end of the porous region, it is reinserted at the
start of the porous region.  To minimize errors caused by reinsertion,
the particle is initialized at the discretization point that has the
closest velocity to the particle's velocity at the outlet.  After a
single trajectory is formed, it has undergone a collection of
reinsertions.  Therefore, as a post-processing step, the trajectory is
made continuous by attaching the tail of each segment to the origin of
the next segment.

@@

The tortuosity definition depends on how the streamlines are
defined, and Duda et al.~\cite{dud-koz-mat2011} summarize several
options. One choice, that we adopt, uses streamlines that originate at
$S$ and terminate at a parallel cross-section.  We use the
cross-sections $x=-1$ and $x=1$ that are at the start and end of the
porous region.  This choice of averaging allows us to also compute the
tortuosity in terms of an area integral.  Assuming that the flow is
incompressible and not re-entrant, meaning that all streamlines connect
the two cross-sections, the tortuosity in
equation~\eqref{eqn:tortuosity1} is equivalent to~\cite{dud-koz-mat2011}
\begin{align}
  T = \frac{\displaystyle\int_\Omega \|\uu(\yy)\|\, d\yy}
           {\displaystyle\int_\Omega u_1(\yy)\, d\yy},
  \label{eqn:tortuosity2}
\end{align}
where $\Omega$ is the fluid region between the inlet and outlet
cross-sections.  Recirculation zones are possible in viscous
fluids~\cite{hig1985}, but they are very small in the examples we
consider and have a negligible effect on the tortuosity.  Since
equation~\eqref{eqn:tortuosity2} does not require the additional work of
computing particle trajectories at every time step, we use this
definition for the majority of the tortuosity calculations.  However, we
do compare the two definitions for the tortuosity at several porosities
in Section~\ref{sec:results}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations}
\label{sec:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by defining the main variables used to model erosion.  We only
briefly summarize the model, and a more detailed description can be
found in previous work~\cite{qua-moo2018}.  We consider flows inside a
confined geometry $\Omega$ that contains $M$ eroding bodies with
boundaries $\gamma_\ell$, $\ell = 1,\ldots,M$.  The boundary of the
fluid domain is $\bd \Omega = \Gamma \cup \gamma_1 \cup \cdots \cup
\gamma_M$, where $\Gamma$ is the outer boundary, taken to be a slightly
smoothed version of $[-3,3] \times [-1,1]$.  We only place eroding
bodies in $[-1,1] \times [-1,1]$ so as to create a buffer region for the
flow profile imposed at the inlet to transition to the more complex flow
intervening between the bodies. On each eroding body, $\nn$ is the
normal vector pointing into the body, and $\ss$ is the unit tangent
vector pointing in the counterclockwise direction.  The normal vector on
$\Gamma$ is the outward unit normal.  Neglecting inertial forces, the
governing equations are
\begin{equation}
\label{eqn:erosionModel}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}, \\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass}, \\
    \uu = \mathbf{0}, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the eroding bodies}, \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity}, \\
    \Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
      &&\mbox{\em erosion model}.
  \end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, $\Vn$ is the normal velocity
of $\gamma$, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on $\gamma$. We simulate erosion by alternating
between solving the fluid equations \nick{and} advancing the \nick{erodible boundaries}. The strength of $\UU$ is adjusted at each time step to achieve a constant pressure drop across
the channel.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Double-Layer Potential Formulation in $\RR^2$}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Applying the same approach as our previous work~\cite{qua-moo2018}, we
write the velocity $\uu$ as a double-layer potential 
\begin{align}
  \DDD[\eeta](\xx) = \int_{\bd\Omega} D(\xx,\yy) \eeta(\yy)\, ds_\yy = 
  \frac{1}{\pi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:velocityDLP}
\end{align}
where $D$ is the kernel of the integral operator, $\rr = \xx - \yy$,
$\rho = \|\rr\|$, $\nn$ is the unit outward normal at $\yy$, and $\eeta$
is an unknown density function.  The double-layer potential \nick{cannot}
represent all solutions of the incompressible Stokes equations, so we
complete the integral equation formulation by adding the $M$ Stokeslets
and rotlets~\cite{pow-mir1987}
\begin{align}
  S[\llambda_\ell,\cc_\ell](\xx) = \frac{1}{4\pi} \left( 
    -\log \rho \II + \frac{\rr \otimes \rr}{\rho^2}
    \right)\llambda_\ell, \quad
  R[\xi_\ell,\cc_\ell](\xx) = \frac{\rr^\perp}{\rho^2} \xi_\ell,
  \quad \ell=1,\ldots,M,
\end{align}
where $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ eroding body, $\rr^\perp = (r_2,-r_1)$, and $\rho =
\|\rr\|$.  Then, for any sufficiently smooth geometry $\Omega$, the
solution of the incompressible Stokes equation with a Dirichlet boundary
condition $\ff$ is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:BIE}
\begin{alignat}{3}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx) +
    \NN_0[\eeta](\xx), \quad &&\xx \in \bd\Omega, \\
  \llambda_\ell &= \frac{1}{2\pi} \int_{\gamma_\ell} 
    \eeta(\yy)\, ds_\yy &&\ell = 1,\ldots,M, \\
  \xi_\ell &= \frac{1}{2\pi} \int_{\gamma_\ell}
    (\yy - \cc_\ell)^\perp \cdot \eeta(\yy)\, ds_\yy 
    &&\ell = 1,\ldots,M.
\end{alignat}
\end{subequations}
Here, the null space associated with the flux-free condition of $\ff$ is
addressed with \nick{\st{the integral operator}} $\NN_0$ which is the integral
operator having kernel $N_0(\xx,\yy) = \nn(\xx) \otimes \nn(\yy)$,
$\xx,\yy \in \Gamma$.  In this work, $\ff$ is the prescribed velocity, \nick{which is equal to  $\UU$ on the outer wall, $\Gamma$, and equal to zero on the eroding bodies,
$\gamma_\ell$, $\ell=1,\ldots,M$.}

Once~\eqref{eqn:BIE} is solved for the density function $\eeta$, the
Stokeslets $\llambda_\ell$, and the rotlets $\xi_\ell$, we are able to
compute the deformation tensor, pressure, and vorticity, defined
respectively as
\begin{align}
  \label{eqn:deformationDLP} 
  \ssigma(\xx) &= \int_{\bd\Omega} \mathbf{D}_{\ssigma}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = \frac{1}{2\pi}\int_{\bd\Omega} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \nonumber \\
    & \qquad\qquad\qquad\qquad\qquad\qquad \left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
    \right) \, ds_\yy, \\
%
  p(\xx) &= \int_{\bd\Omega} D_{p}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\pi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) \nn \cdot
      \eeta(\yy)\,ds_\yy, \\
%
  \label{eqn:vorticityDLP} 
  \omega(\xx) &= \int_{\bd\Omega} D_{\omega}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\pi} \int_{\bd\Omega}
    \frac{(\rr \cdot \nn^\perp) + (\rr \cdot \nn)}{\rho^4} 
      (\rr \cdot \eeta) \, ds_\yy.
\end{align}
Since the deformation tensor of the double-layer potential is
discontinuous across $\bd\Omega$, the jump term 
\begin{align}
  \frac{1}{2} \left(\pderiv{\eeta}{\ss} \cdot \ss \right) \left[
    \begin{array}{cc}
      s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
    \end{array}
  \right],
  \label{eqn:deformationJump}
\end{align}
must be added to $\ssigma(\xx)$ for $\xx \in \bd\Omega$.  Finally, the
Stokeslets and rotlets do not exhibit any jumps across $\gamma$, so the
associated components of the deformation tensor, pressure, and vorticity
due to the Stokeslets and rotlets are straightforward to
compute~\cite{poz1992}.  Having computed the deformation tensor on
$\gamma$, the shear stress is computed using
equation~\eqref{eqn:shearStress}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In equations~\eqref{eqn:velocityDLP} and~\eqref{eqn:deformationDLP},
$\uu$ and its deformation tensor are written as layer potentials in
$\RR^2$, but the quadrature method we introduce in
Section~\ref{sec:method} requires a complex-valued representation of the
layer potentials. The first step to form a complex representation is to
write the Laplace double-layer potential as the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Re (v(x)) \quad 
    \text{where} \quad v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy, \quad x \in \nick{ \CC(\Omega) },
  \label{eqn:laplaceComplex}
\end{align}
where $x,y \in \CC$ are the complex counterparts of $\xx,\yy \in \RR^2$,
and $\eta$ the complex counterpart of $\eeta$. \nick{We use the notation $\CC(\Omega)$ to denote the physical domain $\Omega$ represented in complex coordinates.}
\todo[inline]{What do you think of this notation I introduced to distinguish between $\Omega$ as a real versus complex domain? I was skimming over this section and was very confused about what $x$ and $y$ were, having naively assumed they were physical $x$ and $y$ coordinates instead of complex numbers. Hence I wanted to make their definitions more prominent. I realize that you state $x$ and $y$ are complex numbers, but it is in the text and could be missed by the careless reader (i.e. me).}
Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy integral by
first finding the boundary data of $v$.  Assuming that $\nick{\CC(\Omega)}$ is a
simply-connected interior domain, the boundary data of $v$ satisfies the
Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \nick{\CC(\bd\Omega)}.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
decomposed into its different connected components and the appropriate
jump relation is applied.  Since $v$ is a holomorphic function and we
have its boundary data, by the Cauchy integral theorem we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
\end{align}
for $x \in \nick{\CC(\Omega)}$.  Since $v(x)$ depends on the complex-valued density
function $\eta$, we use the notation $v[\eta](x)$ for the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation for the Gradient of the
Double-Layer Potential}
\label{sec:gradDLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Computing the shear stress and vorticity requires a complex-valued layer
potential representation of the \nick{velocity gradient}.  The
deformation tensor at $x \in \Omega$ is found by computing the
derivatives of the expressions for $u_1$ and $u_2$ in
equation~\eqref{eqn:cauchyVelocity}  
\begin{equation}
\label{eqn:cauchyGradient}
  \begin{aligned}
    \pderiv{u_1}{x_1} &= +\Re (v'[\psi_1](x)) + 
      \Re (v''[y\cdot\eta](x)) - \Re (v'[\eta_1](x)) - 
      x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x)), \\
    \pderiv{u_1}{x_2} &= - \Im (v'[\psi_1](x)) - 
      \Im (v''[y\cdot\eta](x)) + x_1\Im (v''[\eta_1](x)) - 
      \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_1} &= +\Re (v'[\psi_2](x)) - 
      \Im (v''[y\cdot\eta](x)) + \Im (v'[\eta_1](x)) +
      x_1\Im (v''[\eta_1](x)) + x_2\Im (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_2} &= -\Im (v'[\psi_2](x)) - 
      \Re (v''[y\cdot\eta](x)) + x_1\Re (v''[\eta_1](x)) +
      \Im (v'[\eta_2](x)) + x_2\Re (v''[\eta_2](x)).
  \end{aligned}
\end{equation}
The same expressions are used to compute the deformation tensor for $x
\in \bd\Omega$, except that the jump
condition~\eqref{eqn:deformationJump} is included.  Finally, to compute
the shear stress, the deformation tensor on $\bd\Omega$ is applied to
the normal and tangent vectors as in
equation~\eqref{eqn:shearStress}.  The \nick{velocity gradient} is
also used to compute the vorticity in the fluid bulk.  For $x \in
\nick{\CC(\Omega)}$, the Cauchy integral representation of the vorticity is
\begin{align}
  \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} = 
\Re (v'[\psi_2](x)) + \Im (v'[\psi_1](x)) 
 + \Re (v'[\eta_2](x))+ \Im (v'[\eta_1](x)), \quad x \in \Omega.
\end{align}

@@

To establish an anomalous dispersion rate, the trajectories must pass
several grains so that an asymptotic regime is reached. The geometries
that we consider are too short to observe the asymptotic dispersion, so
we use a reinsertion method to form longer trajectories. Similar to the
work of others~\cite{dea-qua-bir-jua2018, puy-gou-den2019}, once a
particle reaches the \nick{outlet} of the porous region, it is reinserted at the \nick{inlet}.  To minimize errors caused by reinsertion, the particle is initialized at the discretization point that has the closest velocity to the particle's velocity at the outlet.  After a
single trajectory is formed, it has undergone a collection of
reinsertions.  Therefore, as a post-processing step, the trajectory is
made continuous by attaching the tail of each segment to the origin of
the next segment.

@@

The tortuosity definition depends on how the streamlines are
defined, and Duda et al.~\cite{dud-koz-mat2011} summarize several
options.  
\todo[inline]{What does it mean `how streamlines are defined'? I had thought that the definition of a streamline is unambiguous. Do you mean which streamlines are used? i.e.~If there is a bias towards either shorter or longer streamlines?}
\nick{\st{One choice, that we adopt, uses }} The choice adopted here is to use
streamlines that originate at $S$ and terminate at a parallel cross-section.  We use the
cross-sections $x=-1$ and $x=1$ that are at the start and end of the
porous region.  This choice of averaging allows us to also compute the
tortuosity in terms of an area integral.  Assuming that the flow is
incompressible and not re-entrant, meaning that all streamlines connect
the two cross-sections, the tortuosity in
equation~\eqref{eqn:tortuosity1} is equivalent to~\cite{dud-koz-mat2011}
\begin{align}
  T = \frac{\displaystyle\int_\Omega \|\uu(\yy)\|\, d\yy}
           {\displaystyle\int_\Omega u_1(\yy)\, d\yy},
  \label{eqn:tortuosity2}
\end{align}
where $\Omega$ is the fluid region between the inlet and outlet
cross-sections.  Recirculation zones are possible in viscous
fluids~\cite{hig1985}, but they are very small in the examples we
consider and have a negligible effect on the tortuosity.  Since
equation~\eqref{eqn:tortuosity2} does not require the additional work of
computing particle trajectories at every time step, we use this
definition for the majority of the tortuosity calculations.  However, we
do compare the two definitions for the tortuosity at several porosities
in Section~\ref{sec:results}.
